\documentclass[12pt]{article}

%\AtEndDocument{\cleardoublepage}
%\AtEndDocument{% based in \cleardoublepage
%  \clearpage
%  \ifodd\value{page}\else
%    \thispagestyle{empty}%
%    \hbox{}\newpage
%  \fi
%}

\usepackage[margin=1.0in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts}
\usepackage{fancyvrb}
\usepackage{graphics}

\usepackage{enumitem}

\theoremstyle{definition}
\newtheorem*{prob}{Problem}
\newtheorem*{soln}{Solution}

\DeclareMathOperator{\lcm}{lcm}

\newcommand{\CC}{{\mathbb{C}}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\RR}{{\mathbb{R}}}
\newcommand{\NN}{{\mathbb{N}}}
\newcommand{\QQ}{{\mathbb{Q}}}
\newcommand{\PP}{{\mathbb{P}}}


\renewcommand{\mod}{\operatorname{mod}}
\renewcommand{\ker}{\operatorname{ker}}

\begin{document}

% T I T L E
%
\title{421 HW 13 Group}
\author{\textbf{Put names here}}

\date{}

\maketitle

\textbf{NOTE}: Unless stated otherwise, $G$ is a (multiplicative) 
group with identity element $e$.



\begin{prob}[8.2.20] \text{ }

(a)   Let $N$ and $K$ be subgroups of a group $G$. 
If $N$ is normal in $G$, prove that $NK = \{nk \mid n\in N, k\in K\}$ 
is a subgroup of $G$. [Compare Exercise 26(b) of Section 7.3.]     

(b) If  both $N$ and $K$ are normal subgroups of  $G$, prove that $NK$ is normal.
\end{prob}

\begin{soln}\quad

(a)

normality: $\forall n\in N,g\in G\quad gng^{-1}\in N$

To prove: $e_G\in NK$

Since $N$ and $K$ are subgroups of $G$, $e_G\in N$, and $e_G\in K$.

Since $e_G\in N$ and $e_G\in K$, $e_Ge_G\in NK$.

$e_G\in NK$

To prove: $NK$ is closed under inverses

Let $nk\in NK$, where $n\in N$ and $k\in K$.

$N$ and $K$ are both groups, so $n^{-1}\in N$ and $k^{-1}\in K$.

Since $n^{-1}\in N$, $N$ is a normal subgroup of $G$, and $k^{-1}\in G$,
$k^{-1}n^{-1}k\in N$.

Since $k^{-1}n^{-1}k\in N$ and $k^{-1}\in K$,
$k^{-1}n^{-1}kk^{-1}\in NK$.

Equivalently, $k^{-1}n^{-1}=(nk)^{-1}\in NK$.

$NK$ is closed under inverses.

To prove: $NK$ is closed under multiplication

Let $nk,mj\in NK$, for some $n,m\in N$ and $k,j\in K$.

Since $N$ is normal, $m\in N$, and $k\in K$, $kmk^{-1}\in K$.

\begin{tabular}{c c c}
$\begin{gathered}
n\in N\quad kmk^{-1}\in N\\
\Downarrow\\
nkmk^{-1}\in N
\end{gathered}$ & &
$\begin{gathered}
k,j\in K\\
\Downarrow\\
kj\in K
\end{gathered}$\\
&
$\begin{gathered}
nkmk^{-1} kj\in NK\\
nkmj\in NK\\
(nk)(mj)\in NK
\end{gathered}$
\end{tabular}

$NK$ is closed under multiplication.

(b)

Let $nk\in NK$, for some $n\in N$ and $k\in K$.

Let $g$ be an arbitrary element of $G$.

Since $N$ and $K$ are normal, $gng^{-1}\in N$ and $gkg^{-1}\in K$.

By the definition of $NK$, $gng^{-1}gkg^{-1}\in NK$.

By inverses, $gnkg^{-1}\in NK$.

$NK$ is normal.

\end{soln}




\begin{prob}[8.3.9]
Let $G = \ZZ_6 \times \ZZ_2$ and let $N$ be the cyclic subgroup $\langle (1, 1) \rangle$. 
Describe the quotient group $G/N$. [That is, what well-known group $G/N$ isomorphic to?]
Justify.
\end{prob}

\begin{soln}

$|G/N|=[G:N]=\frac{|G|}{|N|}=\frac{12}{6}=2$.

All groups of order 2 are isomorphic to $\ZZ_2$.

$G/N$ is isomorphic to $\ZZ_2$.

\end{soln}

\begin{prob}[8.3.25] \text{ }

(a) Find the order of  $\frac{8}{9}$, $\frac{14}{5}$, and 
$\frac{48}{28}$ in the additive group $\QQ/\ZZ$.

(b) Prove that every element of  $\QQ/\ZZ$ has finite order.     

(c) Prove that $\QQ/\ZZ$ contains elements of  every possible finite order.
\end{prob}

\begin{soln}\quad

The identity element is $\ZZ$.

(a)

$|\frac89|=9$, $|\frac{14}5=5$, $|\frac{48}{28}|=|\frac{12}7|=7$.

(b)

Let $\frac ab$ be an arbitrary element of $\QQ/\ZZ$,
for some $a\in \ZZ$ and $b\in\ZZ^+$.

$b\frac ab=a\in\ZZ$, so $|\frac ab|\mid b$.

$|\frac ab|\mid b$, and $b>0$, so $|\frac ab|\leq b$

The order of this element is less than a finite integer, and so must be finite.

Since this element was arbitrary, the order of any element is finite.

(c)

Let $b\in\ZZ^+$.

Consider the element $\frac1b$.

$b\frac1b=1\in\ZZ$, so the order of $\frac1b$ must divide $b$.

For every positive integer $a$ such that $a<b$, $0<a\frac 1b<1$.

This means that $a\frac 1b$ is not an integer, so $\frac1b$
cannot be of order less than $b$.

The only integer which divides $b$ and is not less than $b$ is $b$.

$\frac1b$ must be of order $b$.

Since $b$ was an arbitrary element of $\ZZ^+$,
for every finite order, there exists an element with that order.

\end{soln}

\begin{prob}[8.4.18]
Find all homomorphic images of  $D_4$.
In other words, if $f:D_4 \to H$ is a surjective homomorphism,
then what are all the possibilities for $H$, up to isomorphism? [Hint: 
First Isomorphism Theorem.]
\end{prob}

\begin{soln}

The First Isomorphism Theorem says that $D_4/\ker f\cong H$.

Since $D_4/\ker f\cong H$, $|D_4/\ker f| = |H|$

Theorem 8.13 says that, since $D_4$ is finite, $|D_4/\ker f|=|D_4|/|\ker f|$.

Since $|H|=|D_4|/|\ker f|$, $|H||\ker f|=|D_4$, so $|H|\mid |D_4|$.

Since $|D_4|=8$, $H$ can be of order 8, 4, 2, or 1.

Case $|H|=8$: $H$ and $D_4$ have the same number of elements,
so any surjective homomorphism $f$ is also an injection.
Since $f$ is a surjection, an injection, and a homomorphism,
it is a bijective homomorphism, and therefore an isomorphism.
In this case, $H\cong D_4$.

Case $|H|=4,|\ker f|=2$: There are two possibilities ($\ZZ_4$ and $\ZZ_2\times\ZZ_2$).
Setting the kernel of $f$ to $\{r_0,r_2\}$ makes $D_4/\ker f$
isomorphic to $\ZZ_2\times\ZZ_2$.

By theorem 8.16, the kernel of any $f$ must be a normal subgroup of $D_4$.
If the kernel of $f$ contains $r_1$ or $r_3$, it can't even be a group
since a group must contain both identity and inverses,
and there's just not enough space in a group of order 2.
If the kernel of $f$ contains $d$, $h$, $v$, or $t$,
then (by checking each case), it isn't a normal subgroup.
This excludes $\ZZ_4$ as a possibility.

Case $|H|=2$: There is only one possibility, $\ZZ_2$.

Case $|H|=1$: The only possibility is $\{e\}$.

\end{soln}

\begin{prob}[8.4.26]
Prove that $(\ZZ \times \ZZ)/\langle (2, 2) \rangle \cong \ZZ \times \ZZ_2$. 
[Hint: Show that $f:\ZZ \times \ZZ \to \ZZ \times \ZZ_2$, given by 
$f ((a, b)) = (a - b, [b]_2)$, 
is a surjective homomorphism.]
\end{prob}

\begin{soln}\quad

Let $f:\ZZ \times \ZZ \to \ZZ \times \ZZ_2$ be defined as
$f ((a, b)) = (a - b, [b]_2)$.

To prove: $f$ is surjective

Let $(c, d)$ be an arbitrary element of $\ZZ\times\ZZ_2$.

Let $b$ be $0$ if $d=[0]$, or $1$ if $d=[1]$.

Let $a$ be $c+d$.

$(a, b)$ is an element of $\ZZ\times\ZZ$, and $f((a,b))=(c,d)$.

Since $(c, d)$ was arbitrary, every element in $\ZZ\times\ZZ_2$ has a preimage.

$f$ is surjective.

To prove: $f$ is a homomorphism.

Let $(a, b)$ and $(c, d)$ be arbitrary elements of $\ZZ\times\ZZ$.

To check homomorphism of addition:

\begin{center}
\begin{tabular}{c|c}
$\begin{gathered}
f((a,b))+f((c,d))\\
(a,[b]_2)+(c,[d]_2)\\
(a+c,[b]_2+[d]_2)
\end{gathered}$
&
$\begin{gathered}
f((a,b)+(c,d))\\
f((a+c,b+d))\\
(a+c,[b+d]_2)
\end{gathered}$
\end{tabular}

To check whether $[b]_2+[d]_2=[b+d]_2$, see 4 cases:

\begin{tabular}{c c c c}
$b$ & $d$ & $[b+d]_2$ & $[b]_2+[d]_2$ \\
even & even & 0 & 0\\
even & odd & 1 & 1\\
odd & even & 1 & 1\\
odd & odd & 0 & 0
\end{tabular}

It is always that $[b]_2+[d]_2=[b+d]_2$.

Thus, $(a+c,[b]_2+[d]_2)=(a+c,[b+d]_2)$,

and $f((a,b)+(c,d))=f((a,b))+f((c,d))$.

$f$ is homomorphic under addition.
\end{center}

It is now established that $f$ is a surjective homomorphism.

The kernel of $f$ is all $(a, b)$ for which $a-b=0$ and $[b]_2=0$ -
that is, $a=b$ and $b$ is even.

This is the subgroup of $\ZZ\times\ZZ$ which is generated by $(2,2)$.

The first isomorphism states that,
given groups $G$ and $H$,
and a surjective homomorphism $f: G\to H$,
$G/\ker f\cong H$.

By the first isomorphism,
with $G=\ZZ\times\ZZ$ and $H=\ZZ\times\ZZ_2$,
$\ZZ\times\ZZ/\langle (2,2) \rangle \cong \ZZ\times\ZZ_2$.

\end{soln}


\begin{prob}[3.3.16]
Let $T$, $R$, and $F$ be the four-element rings whose tables are given in Example 5 of  
Section 3.1 and in Exercises 2 and 3 of  Section 3.1. 
Show that no two of  these rings are isomorphic.

For convenience, here are their operation tables:

$T = \{z,r,s,t \}$

\[
\begin{array}{c|cccc}
+ & z & r & s & t \\
\hline
z      & z & r & s & t  \\
r      & r & z & t & s  \\
s      & s & t & z & r  \\
t      & t & s & r & z 
\end{array}
\qquad\qquad
\begin{array}{c|cccc}
\cdot & z & r & s & t \\
\hline
z      & z & z & z & z  \\
r      & z & z & r & r  \\
s      & z & z & s & s  \\
t      & z & z & t & t 
\end{array}
\]

$R = \{ 0, e, b, c \}$

\[
\begin{array}{c|cccc}
+ & 0 & e & b & c \\
\hline
0      & 0 & e & b & c  \\
e      & e & 0 & c & b  \\
b      & b & c & 0 & e  \\
c      & c & b & e & 0 
\end{array}
\qquad\qquad
\begin{array}{c|cccc}
\cdot & 0 & e & b & c \\
\hline
0      & 0 & 0 & 0 & 0  \\
e      & 0 & e & b & c  \\
b      & 0 & b & b & 0  \\
c      & 0 & c & 0 & c 
\end{array}
\]

$F = \{ 0, e, a, b \}$

\[
\begin{array}{c|cccc}
+ & 0 & e & a & b \\
\hline
0      & 0 & e & a & b  \\
e      & e & 0 & b & a  \\
a      & a & b & 0 & e  \\
b      & b & a & e & 0 
\end{array}
\qquad\qquad
\begin{array}{c|cccc}
\cdot & 0 & e & a & b \\
\hline
0      & 0 & 0 & 0 & 0  \\
e      & 0 & e & a & b  \\
a      & 0 & a & b & e  \\
b      & 0 & b & e & a 
\end{array}
\]
\end{prob}

\begin{soln}

They are all isomorphic to $\ZZ_2\times\ZZ_2$ under addition,
so that cannot be used as a basis for finding differences.

$(T,\cdot)$ has an element of order 2 (specifically, $r$),
but $R$ and $F$ do not.
$T$ cannot be isomorphic to $R$ or $F$.

In $R$, there exist two nonzero elements
which multiply to be the additive identity.
This does not happen in $F$,
so these cannot be isomorphic.

\end{soln}

\begin{prob}[3.3.34]
If  $f:R \to S$ is an isomomorphism of  rings, which of  the following properties are 
preserved by this isomorphism? Justify your answers.     

(a) $a\in R$ is a zero divisor.

(b) $a\in R$ is idempotent. (That is, $a^2=a$.)

(c) $R$ is an integral domain.
\end{prob}

\begin{soln}\quad

(a) Yes.

By theorem 3.10.1, $f(0_R)=0_S$.

Lemma: if $b\in R\neq 0_R$, then $f(b)\neq 0_S$.

\leftskip 2cm

Since $f$ is an isomorphism,
$f$ is a bijection,
$f$ is an injection,
and every element of $S$ has at most one preimage.

The preimage of $0_S$ is $0_R$ by theorem 3.10.1,
so no element of $R$ but $0_R$ can map to $0_S$.

\leftskip 0cm

If $a\in R$ is a zero divisor, then

\begin{center}
$\exists b\in R\quad b\neq 0\text{ and } (ab=0_R) \text{ or } ba=0_R)$

\begin{tabular}{c|c}
case $ab=0_R$ & case $ba=0_R$\\
$f(a)f(b)=0_S$ & $f(b)f(a)=0_S$
\end{tabular}
\end{center}

In either case, $f(b)\neq 0_S$ by the lemma a few words ago,
so $f(a)$ is a zero divisor.

(b) Yes.

\begin{align*}
a^2&=a\\
aa&=a\\
f(a)f(a)&=f(a)
\end{align*}

(c) Yes.

Integral domains

\begin{itemize}
\item are commutative
\item have multiplicative identity (which is not equal to additive identity)
\item $ab=0\Rightarrow (a=0\text{ or } b=0)$
\end{itemize}

For commutativity, let $f(a),f(b)$ be any elements of $S$
(which is possible because $f$ is an isomorphism):
\begin{align*}
ab&=ba\\
f(a)f(b)&=f(b)f(a)
\end{align*}

For multiplicative identity,
$R$ has an element $1_R$ such that $\forall r\in R\quad r1_R=1_Rr=r$.
If $f$ is applied, this implies that $S$ has an element $f(1_R)$
such that $\forall f(r)\in S\quad f(r)f(1_R)=f(1_R)f(r)=f(r)$.

Let $f(a),f(b)\in S$ such that $cd=0$.
\begin{align*}
f(a)f(b)&=0_S\\
f(ab)&=0_S\\
ab&=0_R\\
a=0_R&\text{ or }b=0_R\\
f(a)=0_S&\text{ or }f(b)=0_S
\end{align*}
If $R$ has this property, then $S$ has this property.

If $R$ is an integral domain, then $S$ is an integral domain.

\end{soln}

\begin{prob}[3.3.38]
Let $F$ be a field and $f:F \to R$ a homomorphism of  rings.     

(a)   If there is a nonzero element $c$ of  $F$ such that $f(c) = 0_R$, 
prove that $f$ is the zero homorphism (that is, $f(x) = 0_R$ for every $x\in F$). 
[Hint: $c^{-1}$ exists (Why?). If  $x\in F$, consider $f(xcc^{-1})$.]     

(b)   Prove that $f$ is either injective or the zero homomorphism. 
[Hint: If  $f$ is not the zero homomorphism and $f(a) = f(b)$, then $f(a - b) = 0_R$.]
\end{prob}

\begin{soln}

\end{soln}


\begin{prob}[3.3.42]
If  $(m, n) \neq 1$, prove that $\ZZ_{mn}$ is not isomorphic to $\ZZ_m \times \ZZ_n$.
\end{prob}

\begin{soln}

\end{soln}

\end{document}
